Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-05-30 VegOzone.tex
Follow Lisa's suggestions for introduction and move part to discussion

\begin{abstract}
  The unique vegetation of the subarctic region, acclimatized to extremes of cold and midnight sun, is likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}. To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution. This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than sub-Arctic paramterizations. With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition. To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site. We refer to this approach as a 'local climate derived parameterization' of stomatal response functions. With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology. In addition, we evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer. Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August. These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions caused by an extensive blocking event. Such conditions might be more likely in the future under advancing climate change. In contrast, 2019 was an average year. Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species. We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$. For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments. Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
\end{abstract}

@@

Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}. Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}. Hence affecting the global carbon cycle \citep{BGS:Lombardozzi2013} and crop yields. Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone. Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently. Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}. This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}. However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions concerning ozone impacts and air quality \citep{NCC:Lin2020}. Tropospheric background ozone concentrations derived from satellite instruments range between $50-65\,\unit{ppb}$ over Fennoscandia but strongly depend on season \citep{ESA:Cooper2014}.

In the annual cycle, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer. This ozone spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}. \citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.

The precursor gases involved in the formation of ground-level ozone are carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}). In northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}. 
Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil. These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores \citep{Ecosystems:Ghimire2021}, or even attraction of pollinators \citep{TPS:Penuelas2003}. \citet{Ecosystems:Ghimire2021} found no clear evidence that general warming would increase the BVOC emissions in subarctic vegetation, while a reduction of herbivory would decrease them by $44-72\,\%$.

The start of the growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak. At the same time, the growing season is also becoming longer. A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose. For natural and semi-natural vegetation that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future. Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.

Ozone acts as oxidative stress to plants. Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves. Ozone enters the leaves through stomata, leaf pores that enable gas exchange, allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream. Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss. The higher the stomatal conductance, the higher the potential for ozone uptake. Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}. Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}. Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.

To estimate the potential ozone risk on subarctic vegetation, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$). Flux-based metrics, that relate the accumulated dose of ozone to a reduction in biomass, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e. physiology and phenology) as well as key environmental variables that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}. But no bespoke parameterization for subarctic vegetation is available.

The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone. This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage. To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}. Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions. Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}. The exceedance of the CL is calculated by: 
%
\begin{equation}
  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
\end{equation}
%

@@

The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates. Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable. It is important to understand the implications of using these in the subarctic. In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data. This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future. In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  (b)\\
  \includegraphics[width=8.3cm]{javis_funcs_light_hist_grassland}
  %\includegraphics[width=8.3cm]{fig07}
\caption{Construction of bespoke response functions for grassland. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_grassland}
\end{figure}

@@

\begin{table}[t]
  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
    & Cold & 5 & 15 & 100\\
    & Subarctic & 0 & 10 & 100\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
    & Cold & 0 & 15 & 100\\
    & Subarctic & 0 & 10 & 100\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
    & Cold & 5 & 15 & 36\\
    & Subarctic & 0 & 10 & 24\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llcr}
    \tophline
    Species & type & $\alpha$ & $\gamma(0.5)$\\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 0.004 & 165.035\\
    & PPFD0.8 & 0.005 & 132.028\\
    & PPFD1.2 & 0.003 & 198.042\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0.006 & 115.525\\
    & PPFD0.8 & 0.008 & 92.420\\
    & PPFD1.2 & 0.005 & 138.629\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 0.011 & 63.013\\
    & PPFD0.8 & 0.014 & 50.411\\
    & PPFD1.2 & 0.009 & 75.616\\
    \bottomhline
    \end{tabular}
\end{table}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_func_opt_mean}
  %\includegraphics[width=8.3cm]{fig08}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
  \label{fig:javis_func_opt_mean}
\end{figure}

@@

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_deciduous.pdf}\\
  (b)\\
  \includegraphics[width=8.3cm]{javis_funcs_light_hist_deciduous.pdf}
  %\includegraphics[width=8.3cm]{figB4}
\caption{Construction of bespoke response functions for deciduous trees. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_birch}
\end{figure}

>>>>

\begin{abstract}
  The unique vegetation of the subarctic region, acclimatized to extremes of cold and midnight sun, is likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}. To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution. This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than sub-Arctic paramterizations. With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition. To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site. We refer to this approach as a 'local climate derived parameterization' of stomatal response functions. We also study the impact of the opening extent of stomata at low light intensities. With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology. In addition, we evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer. Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August. These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions caused by an extensive blocking event. Such conditions might be more likely in the future under advancing climate change. In contrast, 2019 was an average year. Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species. We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$ compared to $2.3-11.2\,\%$. For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments. Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
\end{abstract}

@@

Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}. Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}. Such ozone induced injury and damage can cause losses in crop \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}. Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites for which such data do exist suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently. Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone concentrations started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}. This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}. However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions on ozone concentrations and associated impacts \citep{NCC:Lin2020}.

Tropospheric background ozone concentrations derived from satellite instruments range between $50-65\,\unit{ppb}$ over Fennoscandia but strongly depend on season \citep{ESA:Cooper2014}. Annual ozone concentration profiles for Fennoscandia show a distinct maximum ground level-ozone concentration during the spring followed by lower concentrations throughout the summer. This ozone spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}. \citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background ozone concentrations determined by circulation patterns and atmospheric transport, whilst summertime abundance is mostly affected by emissions of precursors in the rest of Europe. For example, in northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 were attributed to ozone precursor emissions from forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}. The precursor gases were predominantly carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}). Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil.

The growing season (GS) in northern Fennoscandia has been shifting and starts earlier in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}. Its start is thereby converging with the period of the ozone spring peak. At the same time, the growing season is also becoming longer. A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose. For natural and semi-natural vegetation types that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future. Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.

Ozone acts as oxidative stress to plants. Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves. Ozone enters the leaves through stomata, leaf pores that enable gas exchange, allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream. Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss. The higher the stomatal conductance, the higher the potential for ozone uptake. Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}. Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}. Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.

To estimate the potential ozone risk on subarctic vegetation, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$). Flux-based metrics, that relate the accumulated dose of ozone to a reduction in biomass, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e. physiology and phenology) as well as key environmental variables that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}. Currently, parameterizations are available for some species for particular climatic regions e.g. Boreal, temperate and Mediterranean; there is no parameterization for sub arctic vegetation. Given the unique conditions of the sub arctic (i.e. long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to boreal species.

The $\mathrm{POD_y}$ metric is currently used within the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone. This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage. To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}. Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}. The exceedance of the CL is calculated by: 
%
\begin{equation}
  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
\end{equation}
%

@@

Flux-based methods of risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates. Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that Boreal parameterization will be applicable. In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore a local climate derived temperature response function. We also study the impact of the opening extent of stomata at low light intensities (light response function).\\

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  (b)\\
  \includegraphics[width=8.3cm]{javis_funcs_light_hist_grassland}
  %\includegraphics[width=8.3cm]{fig07}
\caption{Construction of local climate derived response functions for grassland. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_grassland}
\end{figure}

@@

\begin{table}[t]
  \caption{Local climate derived temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
    & Cold & 5 & 15 & 100\\
    & Subarctic & 0 & 10 & 100\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
    & Cold & 0 & 15 & 100\\
    & Subarctic & 0 & 10 & 100\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
    & Cold & 5 & 15 & 36\\
    & Subarctic & 0 & 10 & 24\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Local climate derived light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llcr}
    \tophline
    Species & type & $\alpha$ & $\gamma(0.5)$\\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 0.004 & 165.035\\
    & PPFD0.8 & 0.005 & 132.028\\
    & PPFD1.2 & 0.003 & 198.042\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0.006 & 115.525\\
    & PPFD0.8 & 0.008 & 92.420\\
    & PPFD1.2 & 0.005 & 138.629\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 0.011 & 63.013\\
    & PPFD0.8 & 0.014 & 50.411\\
    & PPFD1.2 & 0.009 & 75.616\\
    \bottomhline
    \end{tabular}
\end{table}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_func_opt_mean}
  %\includegraphics[width=8.3cm]{fig08}
  \caption{Proposed metric to test local climate derived response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
  \label{fig:javis_func_opt_mean}
\end{figure}

@@

..
This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future. In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
..

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_deciduous.pdf}\\
  (b)\\
  \includegraphics[width=8.3cm]{javis_funcs_light_hist_deciduous.pdf}
  %\includegraphics[width=8.3cm]{figB4}
\caption{Construction of local climate derived response functions for deciduous trees. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_birch}
\end{figure}
